Out-of-Sample Comparisons of Overfit Models

### 2012-03-30 paper.tex
Typed up changes to paper

\documentclass[11pt]{article}
%\linespread{1.2}\selectfont
\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{fit}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[round,comma]{natbib}
\usepackage{setspace,amsmath,amssymb,amsthm,ifthen,subfig,url,multicol}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\thepage}\chead{}\rhead{}
\lfoot{}\cfoot{}\rfoot{}
\renewcommand\headrulewidth{0pt}
\onehalfspacing
\usepackage[margin=1in]{geometry}
\usepackage{nath}

@@

\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator*{\plim}{plim}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\atan}{atan*}

\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand{\Citepos}[1]{\Citeauthor{#1}'s \citeyearpar{#1}}

%% from http://blog.blubinc.net/
% Define a command for a given term. This will create a command which gives
% the long name, with a parenthesised acronym, on the first call.
% Additionally it will redefine itself to call the acronym function instead
% on subsequent executions.
%
% Takes:
%	1, command name
%	2, long name
%	3, acronym
% ie \newacronym{ols}{Ordinary Least Squares}{OLS} will define the
% command \ols 
\newcommand{\newacronym}[3]{%
\expandafter\newcommand\expandafter{\csname #1\endcsname}{#2 \textsc{(\MakeLowercase{#3})}%
\expandafter\renewcommand\expandafter{\csname #1\endcsname}{\textsc{\MakeLowercase{#3}}}}}

\newcommand{\oosA}{\bar{D}_a}
\newcommand{\oosB}{\bar{D}_b}

\newcommand{\oost}{\ensuremath{\operatorname{oos-t}}}
\newcommand{\oosSum}[2]{\ensuremath{\sum_{#1=R-\h+#2}^{T-\h}}}

\newcommand{\h}{h}
\newcommand{\InnerBlockL}[1]{\ensuremath{R+(#1-1)b_T + 1}}
\newcommand{\InnerBlockU}[1]{\ensuremath{R+#1 b_T}}
\newcommand{\OuterBlockU}{\lfloor \frac P{b_T} \rfloor}
\newcommand{\SumOuterBlock}[1]{\ensuremath{\sum_{#1=1}^{\OuterBlockU}}}
\newcommand{\SumInnerBlock}[2]{\ensuremath{\sum_{#1=\InnerBlockL{#2}}^{\InnerBlockU{#2}}}}

\newcommand{\CEOuterBlock}[2]{\ensuremath{\E_{\InnerBlockU{(#2-1)}} #1}}

\newcommand{\CenteredAverage}{\frac{\oosA - \E_T\oosT}{\hat\sigma / \sqrt{P}}}
\newcommand{\CenteredAverageI}{\ensuremath{\frac{1}{\hat\sigma
    \sqrt{P}}\oosSum{t}{1} (D_{tR} - \E_R D_{tR})}}
\newcommand{\CenteredAverageII}{\ensuremath{\frac{\sqrt{P}}{\hat\sigma}
    (\E_R\oosA - \E_T\oosB)}}

\newcommand{\MixingaleArray}[1]{\ensuremath{}}

\newcommand{\ZSummand}{D_{sR} - \E_R D_{sR}}
\newcommand{\ZDef}{ \frac1{\sqrt{P}} \SumInnerBlock{s}{i} [\ZSummand]}
\newcommand{\ZSqCE}{\ensuremath{\CEOuterBlock{Z_i^2}{i}}}
\newcommand{\ZTrunc}{\ensuremath{C\sqrt{b_{T}/P}}}
\newcommand{\PScaleTerm}{\ensuremath{\frac{C \sqrt{b_T}}{P}}}

\newcommand{\vWeight}{W((t-s)/\gamma)}
\newcommand{\vSummand}{\ensuremath{(D_{tR} - \E_R D_{tR})(D_{sR} - \E_R D_{sR}) \vWeight}}
\newcommand{\kernelBDefn}[1]{\ensuremath{\frac1{\delta\sqrt{2\pi}}e^{-\frac{#1^2}{2\delta^2}}}}
\newcommand{\kernelB}[1]{\ensuremath{\eta_{\delta}(#1)}}

\newcommand{\vtSum}{\ensuremath{\sum_{t=-P+1}^{2P}}}
\newcommand{\vtSumr}{\ensuremath{\sum_{i=1}^r}}
\newcommand{\vtSuma}{\ensuremath{\sum_{t=(2i-2)b_T-P+1}^{(2i-1)b_T-P}}}
\newcommand{\vtSumb}{\ensuremath{\sum_{t=(2i-1)b_T-P+1}^{2ib_T-P}}}
\newcommand{\vttLower}{\ensuremath{\max(1-t,-b_T)}}
\newcommand{\vttUpper}{\ensuremath{\min(P-t,b_T)}}

\newcommand{\varianceTermIDefn}{\ensuremath{\frac1{P}\oosSum{s,t}{1}
    \vSummand}}
\newcommand{\varianceTermI}{\ensuremath{\sigma_{0\delta}^2}}
\newcommand{\varianceTermII}{\ensuremath{\sigma_{1\delta}^2}}
\newcommand{\varianceTermIIa}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{l=\max(1-t,-P)}^{\min(P-t,P)} (D_{t+l+R,R} - \E_R D_{t+l+R,R})
    W(l/\gamma)}}
\newcommand{\varianceTermIIb}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{j=1-t}^{P-t} (D_{t+j+R,R} - \E_R D_{t+j+R,R})
    \kernelB{j/\gamma}}}
\newcommand{\varianceTermIII}{\ensuremath{\sigma_{2\delta}^{2}}}
\newcommand{\varianceTermIIIa}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{l=\vttLower}^{\vttUpper} (D_{t+l+R,R} - \E_R D_{t+l+R,R})
    W(l/\gamma)}}
\newcommand{\varianceTermIV}{\ensuremath{\sigma_{3\delta}^{2}}}
\newcommand{\varianceTermIVb}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{j=\vttLower}^{\vttUpper} (D_{t+j+R,R} - \E_R D_{t+j+R,R})
    \kernelB{j/\gamma}}}
\newcommand{\vtIIIsummand}{\ensuremath{(Z_{1t}Z_{2t} - \E_RZ_{1t}Z_{2t})}}

\newcommand{\varianceDiffA}{\ensuremath{\frac1{P} \oosSum{s,t}{1}
    \lvert (D_{tR} - \E_R D_{tR})(\E_R D_{sR} - \oosA) \rvert
    \vWeight}}
\newcommand{\varianceDiffAi}{\ensuremath{\frac1P \oosSum{s}{1}(\E_R
    D_{sR} - \oosA )^2}}
\newcommand{\varianceDiffAii}{\ensuremath{\frac1P \oosSum{s}{1}
    (D_{sR} - \E_R D_{sR})^2}}

\newcommand{\rhoExp}{\ensuremath{\frac{\rho-2}{2\rho}}}
\newcommand{\absReg}{\ensuremath{\frac\rho{\rho-2}}}

\newcommand{\couplingConstant}{\ensuremath{2^{\frac{1+\rho}{\rho}} B_L}}
\newcommand{\couplingBeta}[1]{\ensuremath{\beta^{\frac{\rho-2}{2\rho}}_{#1}}}
\newcommand{\couplingBetaSq}[1]{\ensuremath{\beta^{\frac{\rho-2}{\rho}}_{#1}}}
\newcommand{\couplingBound}[1]{\couplingConstant \couplingBeta{#1}}

\newcommand{\OCoef}{\ensuremath{O_p(\frac{\sqrt{P}}{T})}}

\newcommand{\bh}[1]{\ensuremath{\hat\theta_{#1}}}
\newcommand{\ep}[1]{\ensuremath{\boldsymbol{\varepsilon}_{#1}}}

\newcommand{\olsMatrix}[2]{\ensuremath{(#1'#1)^{-1}#1'#2}}

\newcommand{\dmw}{\textsc{dmw}}
\newcommand{\ols}{\textsc{ols}}
\newcommand{\mds}{\textsc{mds}}
\newcommand{\clt}{\textsc{clt}}
\newcommand{\lln}{\textsc{lln}}
\newacronym{VAR}{Vector Autoregression}{VAR}
\newacronym{oos}{out-of-sample}{OOS}
\newcommand{\dgp}{\textsc{dgp}}
\newcommand{\hac}{\textsc{hac}}
\newacronym{ned}{Near Epoch Dependent}{NED}
\newacronym{mse}{Mean Squared Error}{MSE}
\newcommand{\sfield}{$\sigma$-field}
\newcommand{\sfields}{\sfield s}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}

\frenchspacing \title{Out-of-Sample Comparisons of Overfit Models}
\author{Gray Calhoun\thanks{email: \texttt{gcalhoun@iastate.edu}. I
    would like to thank Julian Betts, Helle Bunzel, Marjorie Flavin,
    Nir Jaimovich, Lutz Kilian, Michael McCracken, Seth Pruitt, Ross
    Starr, Yixiao Sun, Allan Timmermann, participants at the 2010
    Midwest Economics Association Annual Meetings, the 2010 Joint
    Statistical Meetings, and in many seminars at \textsc{ucsd}, and
    espeically Graham Elliott for their valuable suggestions, feedback
    and advice in writing this paper.  I would also like to thank Amit
    Goyal for providing computer code and data for his 2008 RFS paper
    with Ivo Welch \citep{GoW:08}.} \\ Iowa State University,
  Economics Department}
\date{October 24, 2011}

@@

\begin{abstract}\thispagestyle{empty}\noindent
  This paper uses dimension asymptotics to study why overfit linear
  regression models should be compared out-of-sample; we let the
  number of predictors used by the larger model increase with the
  number of observations so that their ratio remains uniformly
  positive.  Under this limit theory, the naive Diebold-Mariano-West
  out-of-sample test can test hypotheses about a key quantity for
  evaluating forecasting models---a time series analogue to the
  generalization error---as long as the out-of-sample period is small
  relative to the total sample size.  Moreover, tests that are
  designed to reject if the larger model is true, such as the usual
  in-sample Wald and \textsc{lm} tests and also Clark and McCracken's
  (2001, 2005a), \citepos{Mcc:07} and Clark and West's (2006, 2007)
  out-of-sample statistics, will choose the larger model too often
  when the smaller model is more accurate.

\noindent \textsc{jel} Classification: C01, C12, C22, C52, C53

\noindent Keywords: Generalization Error, Forecasting, Model
Selection, t-test, Dimension Asymptotics
\end{abstract}

\section{Introduction}\label{sec:introduction}
Consider two sequences of length $P$ of prediction errors, the result
of forecasting the same variable with two different estimated models.
Both models are estimated with $R$ observations, collectively called
the {\em estimation window}, and are used to forecast an additional
$P$ observations, called the {\em test sample}.  There are $T$
observations in all, and $R+P=T$.  This paper introduces a new limit
theory for statistics constructed from these prediction errors
designed to approximate the behavior of the statistics when one of the
models is overfit.  In doing so, we provide a theoretical
justification for forecasters to use \oos\ instead of
in-sample comparisons.

Although \oos\ comparisons have been popular in macroeconomics and
finance since \citepos{MeR:83} seminal study of exchange rate models,
it is unclear from a theoretical perspective whether or not the
statistics are useful.  Empirical researchers often cite ``overfit''
or ``instability'' as reasons for using \oos\ comparisons, as in
\citet{StW:03}, but neither term is precisely defined or formalized.
Compounding this problem, the asymptotic distributions of these
statistics are derived under conditions that rule out either
instability or overfit and allow a researcher to use a conventional
in-sample comparison---a variation of the F-test, for example.  As
\citet{InK:04} argue, the statistics themselves are designed to test
hypotheses that can be tested by these in-sample statistics.  For
example, \citet{DiM:95} and \citet{Wes:96} derive the limiting
distributions of many popular \oos\ test statistics under conditions
that would justify these full-sample tests.\footnote{In this paper, we
  will refer to their basic \oos\ $t$-test as the \dmw\ test.}  Much
of the subsequent research by \citet{Mcc:00, Mcc:07}, \citet{CCS:01},
\citet{ClM:01,ClM:05}, \citet{CoS:02,CoS:04}, \citet{ClW:06,ClW:07},
\citet{Ana:07}, and others relaxes several of Diebold and Mariano's
and West's assumptions, but maintains the stationarity and dependence
conditions that permit in-sample comparisons.\footnote{Like us,
  \citet{Ana:07} allows the number of regressors to increase with $T$.
  But in that paper, they increase slowly enough that the \ols\
  coefficients are consistent and asymptotically normal.}
\citet{GiW:06}, and \citet{GiR:09, GiR:10} are exceptions.  Instead of
focusing on hypotheses that can be tested by in-sample comparisons,
\citet{GiW:06} derive an \oos\ test for the null hypothesis that the
difference between two models' \oos\ forecasting performance is
unpredictable, a martingale difference sequence; \citet{GiR:09} test
whether the \oos\ forecasting performance of a model suffers from a
breakdown relative to its in-sample performance; and \citet{GiR:10}
test whether the forecasting performance is stable. However, those
papers focus on a particular \oos\ estimation strategy and do not
address why naive \oos\ comparisons might be generally useful.

@@

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of \oos\ comparisons when at least
one of the models is overfit---when the number of its regressors
increases with the number of observations so that their ratio remains
positive.  We focus on linear regression models estimated with a fixed
training window, but our basic conclusions should be true for other
estimation strategies and models as well.  Under this asymptotic
theory, the \ols\ coefficient estimator is no longer consistent or
asymptotically normal \citep{Hub:73} and has positive variance in the
limit.  Even so, we show that the usual \oos\ average is
asymptotically normal and can consistently estimate the difference
between the models' expected loss in periods $T+1,T+2,\dots$
conditional on the on the data available in period $T$, a time-series
analogue to the generalization error.\footnote{See, for example,
  \citet{HTF:08}.}  Under these asymptotics, the generalization error
does not converge to the expected performance of the pseudo-true
models, so the in-sample and \oos\ comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our theory,
the model that is closer to the true \dgp\ in population can forecast
worse.  In such a situation, an in-sample comparison would correctly
reject the null hypothesis that the benchmark is true, and an \oos\
comparison would correctly fail to reject the null that the benchmark
is more accurate.\footnote{In a pair of papers similar to ours,
  \citet{ClM:09,ClM:09b} study in-sample and \oos\ tests that the
  larger model has nonzero coefficients that are too close to zero to
  expect the model to forecast more accurately.  Like this paper, they
  argue that the larger model can be true but less accurate.  However,
  they focus on an aspect of the \dgp\ that makes this phenomenon
  likely, while we focus on the coefficient estimates that produce
  less accurate forecasts.}

This result justifies \oos\ comparisons when researchers want to
choose a model for forecasting.  Although there has been little
emphasis on hypothesis testing in this setting, testing is usually
appropriate: there is usually a familiar benchmark model in place, and
the cost of incorrectly abandoning the benchmark for a less accurate
alternative model is higher than the cost of incorrectly failing to
switch to a more accurate alternative.\footnote{These costs can be
  monetary, in the form of computer upgrades, software installation,
  and training, or they can be reputational costs.}
We show that the \dmw\ test lets the forecaster control the
probability of the first error, just as with conventional hypothesis
testing.

Since the models' coefficients are imprecisely estimated in the limit,
the test sample must be small to ensure that the model estimated over
the training sample is similar to the one that will be estimated over
the full sample.  In particular, $P/T \to 0$ is required for
consistent estimation, and $P^2/T \to 0$ is required for valid
confidence intervals and inference.  For larger $P$, the \oos\
comparisons remain asymptotically normal, but are centered on the
forecasting performance associated with the period-$R$ estimates.  In
practice, researchers typically use large values of $P$, so these
studies may be too pessimistic about their models' future accuracy if
they use the \dmw\ test.  Section~\ref{sec:oostheory} lays out the
asymptotic behavior of the \dmw\ test under this limit theory.

Since the \ols\ coefficient estimator is no longer consistent or
asymptotically normal, the F-test and Wald tests are no longer
asymptotically valid.  Moreover, \citet{Efr:86,Efr:04} shows that
naive comparisons of in-sample loss overestimate models' accuracy by a
factor proportional to $K/T$, $K$ being the number of regressors.  But
modifications of the F-test are valid under this asymptotic theory, as
shown by \citet{BoB:95}, \citet{AkA:00}, \citet{AkP:04}, and
\citet{Cal:10}, among others.  Such results establish that hypotheses
about the parameters of the \dgp\ can be tested in-sample even if one
or both models are overfit.  For that very reason, these tests can not
reliably choose a model for forecasting and will reject the benchmark
too often.  We show this analytically in Section~\ref{sec:insample}.
Moreover, under this asymptotic theory, many recent \oos\ test
statistics, such as those derived by \cite{ClM:01,ClM:05},
\citet{Mcc:07}, and \citet{ClW:06,ClW:07} should behave like the
F-test.  These tests are also designed to reject the benchmark when
the alternative model is true, and so they may reject too often when
the benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
0$ (with the number of regressors fixed), these statistics behave like
the \dmw\ test when $P$ is small, but should over-reject the benchmark
when $P$ is large.  Section~\ref{sec:mc} presents simulations that
support these claims.

Finally, this paper also introduces a new method of proof for \oos\
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of \oos\ loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales, so asymptotic results for \ned\
functions of mixing processes can be used for these \oos\ processes
with only slight modification.  Although we only present a proof that
this mixingale result holds for the fixed-window scheme, it is simple
to extend it to other estimation windows.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces our notation and assumptions.
Section~\ref{sec:theory} gives the main theoretical results for the
\dmw\ \oos\ test and shows that the F-test rejects the benchmark model
too often when it is false but more accurate than the
alternative. Section~\ref{sec:mc} presents a Monte Carlo study
supporting our theoretical results.  Section~\ref{sec:empirics}
applies the \oos\ statistic to \citepos{GoW:08} dataset
for equity premium prediction, and Section~\ref{sec:conclusion}
concludes.  Proofs and supporting results are listed in the Appendix.

\section{Setup and Assumptions}\label{sec:assumptions}
We will set up our models and notation first, then list the assumptions
required for our results.  There are two competing linear models that
give forecasts for the target, $y_{t+\h}$:
\[
y_{t+\h} = x_{1t}'\theta_1 + \varepsilon_{1,t+h}
\]
and
\[
y_{t+\h} = x_{2t}'\theta_2 + \varepsilon_{2,t+h};
\]
$\h$ is the forecast horizon, and the variables $y_t$, $x_{1t}$, and
$x_{2t}$ are all known in period $t$.  Let
\begin{equation*}
  \mathcal{F}_t = \sigma(y_1, x_1, \dots, y_t, x_t)
\end{equation*}
with $x_t$ the vector of all elements of $x_{1t}$ and $x_{2t}$ after
removing duplicates, and let $\E_t$ and $\var_t$ denote the
conditional mean and variance given $\mathcal{F}_t$.  The first model
uses $K_1$ regressors, and the second uses $K_2$.  Without loss of
generality, assume that $K_1 \leq K_2$.  We will allow $K_1$ and $K_2$
to vary with $T$, so a stochastic array underlies all of our
asymptotic theory, but we suppress that notation to make the
presentation more clear.

Although our theory can apply to many different \oos\ statistics,
we'll only present results for the \dmw\ \oos\ $t$-test.  Both models
are estimated using \ols\ over the training sample:
\[
\hat y_{i,t+\h} = x_{it}'\bh{iR}, \quad i=1,2, \quad t = R+1,\dots,T-h
\]
with
\[
\bh{it} = (\sum_{s=1}^{t-\h} x_{is}x_{is}')^{-1} \sum_{s=1}^{t-\h} x_{is} y_{s+\h},
\quad i=1,2.
\]
The \oos\ average is denoted $\oosA$:
\[
\oosA \equiv P^{-1} \oosSum{t}{1} D_{tR}
\quad \text{with} \quad D_{tS} \equiv L(y_{t+\h} - x_{1t}'\bh{1S}) -
L(y_{t+\h} - x_{2t}'\bh{2S}),
\]
$P = T - R$ and $L$ a known loss function.  The \oos\ statistic
of interest is 
\[
\oost \equiv \frac{\sqrt{P} \oosA}{\hat\sigma},
\]
for $\hat\sigma^2$ some (possibly Heteroskedasticity- and
Autocorrelation-Consistent, or \hac) estimator of the asymptotic
variance of $\oosA$.

>>>>

\documentclass[11pt]{article}
\usepackage{tikz,setspace}
\usetikzlibrary{calc}
\usetikzlibrary{fit}
\usepackage[round]{natbib}
\usepackage[margin=1in]{geometry}
\onehalfspacing

\input{macros}

@@

\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand{\Citepos}[1]{\Citeauthor{#1}'s \citeyearpar{#1}}
\newcommand{\clws}{\citeauthor{ClW:06}'s \citeyearpar{ClW:06,ClW:07}}

%% from http://blog.blubinc.net/
% Define a command for a given term. This will create a command which gives
% the long name, with a parenthesised acronym, on the first call.
% Additionally it will redefine itself to call the acronym function instead
% on subsequent executions.
%
% Takes:
%	1, command name
%	2, long name
%	3, acronym
% ie \newacronym{ols}{Ordinary Least Squares}{OLS} will define the
% command \ols

\title{Out-of-Sample Comparisons of Overfit Models}
\author{Gray Calhoun\thanks{email: \texttt{gcalhoun@iastate.edu}. I
    would like to thank Julian Betts, Helle Bunzel, Marjorie Flavin,
    Nir Jaimovich, Lutz Kilian, Ivana Komunjer, Michael McCracken,
    Seth Pruitt, Ross Starr, Jim Stock, Yixiao Sun, Allan Timmermann,
    Hal White, several anonymous referees, participants at the 2010
    Midwest Economics Association Annual Meetings, the 2010
    International Symposium on Forecasting, the 2010 Joint Statistical
    Meetings, the 2011 \textsc{nber}/\textsc{nsf} Time-Series
    Conference, and in many seminars at \textsc{ucsd}, and espeically
    Graham Elliott for their valuable suggestions, feedback and advice
    in writing this paper.  I would also like to thank Amit Goyal for
    providing computer code and data for his 2008 \textsc{rfs} paper
    with Ivo Welch \citep{GoW:08}.} \\ Iowa State University}
\date{October 24, 2011}

@@

\begin{abstract}\thispagestyle{empty}\noindent
  This paper uses dimension asymptotics to study why overfit linear
  regression models should be compared out-of-sample; we let the
  number of predictors used by the larger model increase with the
  number of observations so that their ratio remains uniformly
  positive.  Even under this limit theory, the naive
  Diebold-Mariano-West out-of-sample test can test hypotheses about
  the models' generalization error, a key quantity for evaluating
  forecasting models.  Moreover, tests that are designed to reject if
  the larger model is true, such as the usual in-sample Wald and
  \textsc{lm} tests and also Clark and McCracken's (2001, 2005a),
  \citepos{Mcc:07}, and Clark and West's (2006, 2007) out-of-sample
  statistics, will choose the larger model too often when the smaller
  model is more accurate. Similar results hold for statistics used for
  in-sample model selection, such as the \aic, \bic, and adjusted
  $R$-square.

\noindent \textsc{jel} Classification: C12, C22, C52, C53

\noindent Keywords: Generalization Error, Forecasting, Model
Selection, $t$-test, Dimension Asymptotics
\end{abstract}
\newpage

\section{Introduction}\label{sec:introduction}
Consider two sequences of length $P$ of prediction errors, the result
of forecasting the same variable with two different estimated models.
Both models are estimated with $R$ observations, collectively called
the {\em estimation window}, and are used to forecast an additional
$P$ observations, called the {\em test sample}.  There are $T$
observations in all, and $R+P=T$.  This paper introduces a new limit
theory for statistics constructed from these prediction errors
designed to approximate the behavior of the statistics when one of the
models is overfit.  In doing so, we provide a theoretical
justification for forecasters to use \oos\ instead of in-sample
comparisons and discover a practical limitation of existing statistics.

Although \oos\ comparisons have been popular in Macroeconomics and
Finance since \citepos{MeR:83} seminal study of exchange rate models,
it is unclear from a theoretical perspective whether or not the
statistics are useful.  Empirical researchers often cite ``overfit''
or ``instability'' as reasons for using \oos\ comparisons, as in
\citet{StW:03}, but neither term is precisely defined or formalized.
Compounding this problem, the asymptotic distributions of these
statistics are derived under conditions that rule out either
instability or overfit and allow a researcher to use a conventional
in-sample comparison---a variation of the $F$-test, for example.  As
\citet{InK:04} argue, the statistics themselves are designed to test
hypotheses that can be tested by these in-sample statistics.  For
example, \citet{DiM:95} and \citet{Wes:96} derive the limiting
distributions of many popular \oos\ test statistics under conditions
that would justify these full-sample tests.\footnote{In this paper, we
  will refer to their basic \oos\ $t$-test as the \dmw\ test.}  Much
of the subsequent research by \citet{Mcc:00, Mcc:07}, \citet{CCS:01},
\citet{ClM:01,ClM:05}, \citet{CoS:02,CoS:04}, \citet{ClW:06,ClW:07},
\citet{Ana:07}, and others relaxes several of Diebold and Mariano's
and West's assumptions, but maintains the stationarity and dependence
conditions that permit in-sample comparisons \citep[see][for a recent
review of this literature]{Wes:06}.\footnote{Like us, \citet{Ana:07}
  allows the number of regressors to increase with $T$.  But in that
  paper, they increase slowly enough that the \ols\ coefficients are
  consistent and asymptotically normal.}  \citet{GiW:06} and
\citet{GiR:09, GiR:10} are exceptions.  Instead of focusing on
hypotheses that can be tested by in-sample comparisons, \citet{GiW:06}
derive an \oos\ test for the null hypothesis that the difference
between two models' \oos\ forecasting performance is unpredictable, a
martingale difference sequence; \citet{GiR:09} test whether the \oos\
forecasting performance of a model suffers from a breakdown relative
to its in-sample performance; and \citet{GiR:10} test whether the
forecasting performance is stable. However, those papers focus on a
particular \oos\ estimation strategy and do not address why \oos\
comparisons might be useful as a general strategy.

@@

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of \oos\ comparisons when at least
one of the models is overfit---when the number of its regressors
increases with the number of observations so that their ratio remains
positive.  Although overfit is sometimes used to describe the
situation where a forecaster chooses from many different models,
i.e. \textit{data-mining} or \textit{data-snooping}, we view these as
separate issues.  Procedures that account for the presence of many
models have been, and are continuing to be, developed \citep[see, for
example,][]{Whi:00,Han:05,RoW:05,HHK:10,ClM:11}, but it is unclear
whether those procedures should themselves use in-sample or \oos\
comparisons.  Understanding the difference between in-sample and \oos\
comparisons in the context of a simple comparison between two models
is necessary before resolving any new issues that arise with multiple
comparisons.\footnote{Moreover, the empirical research that motivates
  this paper uses \textit{pseudo} \oos\ comparisons and not
  \textit{true} \oos\ comparisons.  Even if a true \oos\ comparison
  could account for some forms of data-snooping better than
  \citepos{Whi:00} \textsc{brc} or its extensions, in-sample and
  pseudo \oos\ comparisons would both be affected by the data-snooping.
  This point is also made by \citet{InK:04}.}

We focus on linear regression models for simplicity, but our basic
conclusions should be true for other models as well.  Under this
asymptotic theory, the \ols\ coefficient estimator is no longer
consistent or asymptotically normal \citep{Hub:73} and has positive
variance in the limit.  Even so, we show that the usual \oos\ average
is asymptotically normal and can consistently estimate the difference
between the models' generalization error, the expected loss in the
future conditional on the on the data available in period
$T$.\footnote{See, for example, \citet{HTF:08} for a discussion of
  generalization error.}  Under these asymptotics, the generalization
error does not converge to the expected performance of the pseudo-true
models, so existing in-sample and \oos\ comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our theory,
the model that is closer to the true \dgp\ in population can forecast
worse.  In such a situation, a standard in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an \oos\ comparison would correctly fail to reject the null that the
benchmark is more accurate.\footnote{In a pair of papers similar to
  ours, \citet{ClM:09b,ClM:12} study in-sample and \oos\ tests that
  the larger model has nonzero coefficients that are too close to zero
  to expect the model to forecast more accurately.  Like this paper,
  they argue that the larger model can be true but less accurate.
  However, they focus on an aspect of the \dgp\ that makes this
  phenomenon likely, while we focus on the coefficient estimates that
  produce less accurate forecasts.  Moreover, the implications of our
  asymptotic theories are different and their paper does not provide a
  reason to do \oos\ comparisons.}

This result justifies \oos\ comparisons when researchers want to
choose a model for forecasting.  Although there has been little
emphasis on hypothesis testing in this setting, testing is usually
appropriate: there is usually a familiar benchmark model in place, and
the cost of incorrectly abandoning the benchmark for a less accurate
alternative model is higher than the cost of incorrectly failing to
switch to a more accurate alternative.  We show that the \dmw\ test
lets the forecaster control the probability of the first error, just
as with conventional hypothesis testing.

Since the models' coefficients are imprecisely estimated in the limit,
the test sample must be small enough that the model estimated over the
training sample is similar to the one that will be estimated over the
full sample.  In particular, $P/T \to 0$ is required for consistent
estimation, and $P^2/T \to 0$ is required for valid confidence
intervals and inference.  For larger $P$, the \oos\ comparisons remain
asymptotically normal, but are centered on the forecasting performance
associated with the period-$R$ estimates.  In practice, researchers
typically use large values of $P$, so these studies may be too
pessimistic about their models' future accuracy if they use the \dmw\
test.  Section~\ref{sec:oostheory} lays out the asymptotic behavior of
the \dmw\ test under this limit theory.

Since the \ols\ coefficient estimator is no longer consistent or
asymptotically normal, the $F$-test and Wald tests are no longer
asymptotically valid.  Moreover, \citet{Efr:86,Efr:04} shows that
naive comparisons of in-sample loss overestimate models' accuracy by a
factor proportional to $K/T$; $K$ is the number of regressors.  But
modifications of the $F$-test are valid under this asymptotic theory,
as shown by \citet{BoB:95}, \citet{AkA:00}, \citet{AkP:04},
\citet{Cal:11c}, and \citet{Ana:12}, among others.  These results
establish that hypotheses about the parameters of the \dgp\ can be
tested in-sample even if one or both models are overfit.  For that
very reason, these tests can not reliably choose a model for
forecasting and will reject the benchmark too often.  We show this
analytically in Section~\ref{sec:insample} and present analogous
results for model selection.  Moreover, under this asymptotic theory,
many recent \oos\ test statistics, such as those derived by
\cite{ClM:01,ClM:05}, \citet{Mcc:07}, and \citet{ClW:06,ClW:07} should
have the same problems as in-sample tests.\footnote{Our theoretical
  results apply directly to \citepos{Mcc:07} \oos\ $t$-test, since it
  simply proposes more liberal critical values for the same test
  statistic that we study.  Since \citet{ClW:06,ClW:07} use a finite
  length estimation window, our asymptotics are incompatible with
  theirs and prevent us from studying their test directly, as well as
  \citepos{GiW:06} and other tests based on \citepos{GiW:06}
  asymptotics.  But \clws\ test can be viewed as a stochastic
  adjustment to the critical values of the usual \oos\ $t$ test, so
  our conclusions should apply informally as well.  Specifically, we
  show that the \dmw\ test rejects with probability equal to nominal
  size when the estimated benchmark model is expected to be more
  accurate, so more liberal critical values result in overrejection.}
These tests are also designed to reject the benchmark when the
alternative model is true, and so they may reject too often when the
benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
0$ (with the number of regressors fixed), these statistics behave like
the \dmw\ test when $P$ is small, but should overreject the benchmark
when $P$ is large.  Section~\ref{sec:mc} presents simulations that
support these claims.

Finally, this paper also introduces a new method of proof for \oos\
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of \oos\ loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales, so asymptotic results for \ned\
functions of mixing processes can be used for these \oos\ processes
with only slight modification.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces our notation and assumptions.
Section~\ref{sec:theory} gives the main theoretical results for the
\dmw\ \oos\ test, shows that standard in-sample tests reject the
benchmark model too often when it is false but more accurate than the
alternative, and shows that standard model selection methods choose the
larger model too often under similar conditions. Section~\ref{sec:mc}
presents a Monte Carlo study supporting our theoretical results.
Section~\ref{sec:empirics} applies the \oos\ statistic to
\citepos{GoW:08} dataset for equity premium prediction, and
Section~\ref{sec:conclusion} concludes.  Proofs and supporting results
are listed in the Appendix.

\section{Setup and Assumptions}\label{sec:assumptions}
We will set up our models and notation first, then list the assumptions
required for our results.

\subsection{Notation and Forecasting Environment}
We assume the following forecasting environment. There are two
competing linear models that give forecasts for the target,
$y_{t+\h}$:
\[
y_{t+\h} = x_{1t}'\theta_1 + \varepsilon_{1,t+h}, \quad\text{and}\quad
y_{t+\h} = x_{2t}'\theta_2 + \varepsilon_{2,t+h};
\]
$t = 1,\dots,T-h$, $\h$ is the forecast horizon, and the
variables $y_t$, $x_{1t}$, and $x_{2t}$ are all known in period $t$.
Let
\begin{equation*}
  \mathcal{F}_t = \sigma(y_1, x_1, \dots, y_t, x_t)
\end{equation*}
with $x_t$ the vector of all elements of $x_{1t}$ and $x_{2t}$ after
removing duplicates, and let $\E_t$ and $\var_t$ denote the
conditional mean and variance given $\mathcal{F}_t$.  The first model
uses $K_1$ regressors, and the second uses $K_2$.  Without loss of
generality, assume that $K_1 \leq K_2$.  We will allow $K_1$ and $K_2$
to vary with $T$, so a stochastic array underlies all of our
asymptotic theory, but we suppress that notation to simplify the
presentation.
